The Slowdown of TFP \\ Exogenous and Endogenous Mechanisms

### 2018-01-19 Latex/Dissertation_Workshop_23Jan2018/dw23jan2018.tex
dissertation workshop presentation and a workspace with nsimul =500 and right TFP magnitude, as well as new figs

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A very big question}

Broad consensus on secular stagnation.

@@

The literature has agreed that there are two main factors driving this:
\begin{itemize}
\item decreasing labor force participation
\item decreasing TFP growth
\end{itemize}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A smaller, but still big question}

	Why has TFP slowed down?

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Two possible mechanisms}

\begin{itemize}
\item a permanent, or very persistent exogenous shock to TFP

\item an endogenous mechanism
\end{itemize}

@@

$\rightarrow$ What we do in a nutshell: test for the endogenous mechanism (main analysis will be a SVAR)

@@

@@

\hyperlink{BS_quote}{\beamergotobutton{Barsky \& Sims quote}}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Why more convincing?}
	\label{convincing}

	Because the endogenous TFP literature faces a big problem: The endogenous mechanism is rationalized entirely using R\&D investment.

@@

\begin{enumerate}
\item  But R\&D in the data is almost acyclical.... $\rightarrow$ hard to rationalize as a driver of business cycle fluctuations % hyperlink to graph: growth rate of RD vs. growth rate of TFP and GDP

@@

\item Timing issue: the TFP slowdown begins around 2006, i.e. \emph{before} the Great Recession and thus \emph{before} the marked drop in R\&D  % hyperlink to graph: level of RD and TFP, so we see that TFP drops in 2006, while RD only during the recession.
\hyperlink{timing}{\beamergotobutton{Graph of R\&D}}

@@

$\Rightarrow$ Following the suggestion of Fernald, Hall, Stock \& Watson (2009), we propose to use investment in information technology (IT)  % hyperlink to our own graph motivation.pdf: the top graph.
\hyperlink{it_investment}{\beamergotobutton{Graph of IT investment}}

@@

\begin{itemize}
\item $RP = \pi^{IT}/\pi^{CPI}$. 
\item All variables are real (except price indexes) and in log levels (except for RP, which is in growth rates). 
\item The dataset ranges from 1989:q1 - 2017:q1.
\end{itemize}

@@

$D$ will be a tool to impose our identification assumptions \hyperlink{Technicalities}{\beamergotobutton{Technicalities}}

%%%%%%% Slide %%%%%%
\begin{frame}
\frametitle{Identified shocks}

\begin{itemize}
	\item We impose restrictions on $D$ in order to identify two shocks:
	\begin{enumerate}
		\item News Shock
		\item IT Productivity Shock
	\end{enumerate}

	\item Econometric challenge is to disentangle two shocks with very similar features 
		\begin{enumerate}
		\item No impact effect on TFP
		\item Persistent positive effect over time on TFP
		\item Most likely both shocks have positive impact effects on forward-looking variables
	\end{enumerate}

	\item Barsky \& Sims' identification strategy is not sufficient to disentangle the two
	\begin{itemize}
		\item We need a further assumption (restriction) to find a dimension of difference between the shocks
	\end{itemize}
\end{itemize}

%%%%%%% Slide %%%%%%
\begin{frame}
\frametitle{Additional restriction}

\begin{itemize}
	\item We rely on simple demand and supply theory

	\begin{enumerate}
		\item The IT shock is a \emph{sectoral} shock $\rightarrow$ we expect it to move relative prices
		\item The news shock is \emph{not} a sectoral shock $\rightarrow$ we have no a priori sense of what it should do to relative prices on impact, but after some time relative prices should go back to their initial value
	\end{enumerate}

	$\rightarrow$ add the restriction that a news shock should have no effect on relative prices after a reasonable time

\hspace{4cm} $\hookrightarrow$ enough for prices to adjust, so

\hspace{4cm}   between 6-12 quarters

@@

\begin{equation}
TFP_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{IT_{t-k}}_\text{IT shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation}

@@

    \begin{enumerate}
    	\item The news shock $V_{t-k}$ maximizes the FEV of future TFP subject to the restriction that it has no effect on the relative price $RP$ at a small number of quarters;

@@

	\begin{itemize}
	\item Recall: dataset is quarterly and covers 1989:q1-2017-q1.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{TFP response to both shocks}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_News_Shock_on_TFP__}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_IT_Shock_on_TFP__}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

	\begin{itemize}
	\item both the Barsky \& Sims result ...

@@

	\item ... as well as the the conjecture that the IT shock looks similar to news along certain dimensions,

@@

	\item ... but relative prices do indeed introduce a margin of difference between the two shocks.
	\end{itemize}

@@

\item Shares of FEV of TFP explained ...
	\begin{itemize}
	\item ... are also in line with the Barsky \& Sims result ...
	\item [] For BS, news explains around 45\%, compared to 20\% here

@@

	\item ... yet suggest that the IT shock plays an important role as well (around 52\%)

@@

	\item ... and indeed the IT shock \emph{complements} the news shock, instead of substituting for it
	\item [] For BS, the single identified shock explains around 45\%, while our two identified shocks explain around 73\%
	\end{itemize}

@@

@@

\begin{itemize}
\item Provided what we think (and we hope you do too) a careful and more general test for the endogenous mechanism in TFP.
\begin{itemize}
\item While the literature focuses exclusively on R\&D, we show that IT is a better variable to focus on.
\end{itemize}

@@

\item The results show that by controlling for the presence of news shocks, the endogenous component of TFP is an important driver of TFP fluctuations at long horizons.

@@

@@

\begin{itemize}
\item Rationalize this setting using a theoretical model
\item [] $\hookrightarrow$ what we have in mind: an endogenous TFP model (in the vein of Comin \& Gertler) augmented with news shocks

@@

\item Estimate the model (IR-matching)

@@

\item Perform a counterfactual for the Great Recession \emph{and} the period prior, shutting down the negative IT shocks around 2000
% assuming the IT bubble, i.e. the negative shock to IT productivity around 2000, did not happen.

@@

\item See if we can find interesting interactions between the two shocks?
\item [] $\hookrightarrow$ In particular, we're thinking of noise shocks on IT productivity
	\begin{itemize}
	\item [] news shock: shock to expected future value
	\item [] noise shock: shock to expected current value (``perceptions'')
	\end{itemize}
\end{itemize}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Barsky \& Sims say:}
	\label{BS_quote}

@@

\hyperlink{related_lit}{\beamerreturnbutton{Return}}	
\hyperlink{BS_FEV}{\beamerbutton{Barsky \& Sims' FEV-table}}	
\end{frame}
%%%%%%%%%%%%%%%%%

\vspace{-1cm}
\noindent
\begin{figure}
\centering
\includegraphics[scale=0.3]{fig_BS_fevtable}
\end{figure}

>>>>

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation}

	\begin{itemize}
	\item The neoclassical tradition (RBC literature) in macro has treated total factor productivity (TFP) as exogenous.

	(King \& Rebelo 1999, Christiano, Eichenbaum \& Trabandt 2014, 2018)

@@

	\item The Great Recession 2008 revived interest in endogenous components in TFP for explaining long-run TFP fluctuations.

	\item[] $\rightarrow$ We explore the possibility that general-purpose technologies (GPT), in particular for the last 30 years information technologies (IT), play a role in explaining TFP fluctuations at long horizons.

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Motivation II}

	\begin{itemize}
	\item Our main analysis is estimating a structural VAR in which we identify a shock to the exogenous component of future TFP (a news shock) and a shock to the endogenous component of TFP (a shock to IT productivity, hereafter ``IT shock").

@@

	\item Along the way, we provide a solution to an econometric challenge in the literature on long-run productivity.

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Our starting point: Barsky \& Sims (2011)}
	\label{BS_quote}

BS identify news shocks as shocks that maximize future fluctuations in TFP. But they warn:

$\rightarrow$ We focus on a specific subset of shocks which are ``not really `news''': shocks to the productivity of the sector that produces IT goods.

%\hyperlink{BS_FEV}{\beamerbutton{Barsky \& Sims' FEV-table}}	
\end{frame}
%%%%%%%%%%%%%%%%%

@@

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A toy model}

	How should we think of this IT productivity shock?

@@

@@

@@

	Contextualize using a simple expanding variety model (Romer 1990)

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Expanding variety model}
	\label{expanding_variety}

Final good $Y_t$ is the CES aggregate of a number $A_t$ intermediate goods $Y_t^M(s)$:	% theta is between 0 and infinity
	\begin{equation}
	Y_t = \Big[ \int_{0}^{A_t} Y_t^M(s)^{\frac{\theta-1}{\theta} } ds \Big]^{\frac{\theta}{\theta-1}}
	\end{equation}

\pause

Intermediate good $s$ is Cobb-Douglas with exogenous TFP $\Psi_t$:
\begin{equation}
Y_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
\end{equation}

\pause

$\rightarrow$ letting $K_t$ and $L_t$ be aggregate capital and labor, final output is:
\begin{equation}
Y_t = A_t^{\frac{1}{\theta-1}} \Psi_t K_t^{\alpha}L_t^{1-\alpha}
\end{equation}

\pause

The IT sector is a separate sector that uses IT investment $S_{i,t}$ to expand $A_t$. Importantly, we assume that the IT sector is also augmented by $\Psi_t$, as well as by the sector-specific productivity $\lambda_t$.

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t}) 
\end{equation}

\hyperlink{it_sector}{\beamergotobutton{IT sector in detail}}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{The shocks}

\begin{equation*}
Y_t \; = \underbrace{A_t^{\frac{1}{\theta-1}}}_\text{endogenous TFP} \; \; \underbrace{\Psi_t}_\text{exogenous TFP} \; K_t^{\alpha}L_t^{1-\alpha}
\end{equation*}

$\hookrightarrow$ A news shock is $\Psi_{t+k} \uparrow$ for some $k>0$

$\hookrightarrow$ An IT productivity shock is $\lambda_t \uparrow$, which shows up in $A_{t+j} \uparrow$ with a lag $j>0$

Crucially, both lead to an increase in future TFP, with no effect on impact.

 $\rightarrow$ to a Barsky \& Sims-type identification, the two shocks are observationally equivalent!

 $\Rightarrow$ econometric challenge is to disentangle the two!

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Identification: relative prices}

The price level $P$ is an aggregate of the prices intermediate goods. These are determined as an FOC to 
\begin{equation*}
Y_t(s) = \Psi_t K_t(s)^{\alpha}L_t(s)^{1-\alpha}
\end{equation*}

\hspace{5cm} $\hookrightarrow P$ is a function of $\Psi_t$

The price of IT good, $P^{IT}$, is determined as an FOC to

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t})
\end{equation}

\hspace{5cm}  $\hookrightarrow P^{IT}$ is a function of $\Psi_t$ \textbf{and} $\lambda_t$

$\Rightarrow$ The two shocks imply different responses of relative prices $\frac{P^{IT}}{P}$!

\begin{itemize}
\item A news shock moves both $P$ and $P^{IT}$, leaving relative prices unchanged.
\item An IT productivity shock moves only $P^{IT}$, thus changing relative prices.
\end{itemize}

@@

\begin{itemize}
\item $RP = \pi^{IT}/\pi^{CPI}$. 
\item All variables are real (except price indexes) and in log levels (except for RP, which is in growth rates). 
\item The dataset ranges from 1989:q1 - 2017:q2.
\end{itemize}

@@

We impose our identifying assumptions on the matrix $D$. \hyperlink{Technicalities}{\beamergotobutton{Technicalities}}

@@

\begin{equation}
\epsilon^{TFP}_t =   \underbrace{V_{t-k}}_\text{news shock} + \underbrace{IT_{t-k}}_\text{IT productivity shock}  + \underbrace{\varepsilon_t}_\text{surprise tech shock} 
\end{equation}

@@

    \begin{enumerate}
    	\item The news shock $V_{t-k}$ maximizes the FEV of future TFP subject to the restriction that it has no effect on the relative price $RP$ at a small number of quarters (enough for prices to adjust, so 6-12);

@@

	\begin{itemize}
	\item Recall: dataset is quarterly and covers 1989:q1-2017-q2.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame}
\frametitle{TFP response to both shocks}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_News_Shock_on_TFP__} % REGENERATE TFP SHOCKS FIGURES WITH RIGHT SCALES!!
\end{figure}
\begin{figure}
	\centering
	\includegraphics[scale=0.15]{\ourFigPath Figures/fig_IT_Shock_on_TFP__}
\end{figure}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

	\begin{itemize}
	\item both the Barsky \& Sims result;

@@

	\item as well as the the conjecture that the IT shock looks similar to news along certain dimensions;

@@

	\item but relative prices do indeed introduce a margin of difference between the two shocks.
	\end{itemize}

@@

\item Shares of FEV of TFP explained ...
	\begin{itemize}
	\item are also in line with the Barsky \& Sims result;
	\item [] (for BS, news explains around 45\%, compared to 20\% here)

@@

	\item  yet suggest that the IT shock plays an important role as well (around 52\%).

@@

	\item And indeed the IT shock \emph{complements} the news shock, instead of substituting for it.
	\item [] For BS, the single identified shock explains around 45\%, while our two identified shocks explain around 73\%.
	\end{itemize}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{A counterfactual}

	\begin{figure}
		\centering
		\includegraphics[scale=0.28]{\ourFigPath Figures/fig_counterfactual_}
	\end{figure}

@@

\begin{itemize}
\item Provided a test for the role of IT as an example of GPTs in explaining fluctuations in long-run TFP, and in doing so, overcame an econometric challenge prevalent in the literature on long-run productivity.

@@

\item The results show that by controlling for the presence of news shocks, IT productivity shocks are important drivers of TFP fluctuations at long horizons.

@@

\item Moreover, IT productivity shocks can be thought of as giving more microfoundations to what classical news shocks carry information on.

@@

\begin{itemize}

\item Technically, should do a VECM due to cointegration and wanting to have variables in the VAR as growth rates rather than levels...

@@

\item Dig deeper: IT is just an example of GPTs for the last 30 years
\item[] $\hookrightarrow$ could redo analysis for different time periods with different GPTs (electricity in 1920s, airplane industry in 1960s ...)

@@

\item Use a fully structural model to see if we can find interesting interactions between the two shocks?
\item [] $\hookrightarrow$ In particular, we're thinking of noise shocks on IT productivity

@@

\item But there are also other shocks that are similar to news shocks and yet not news shocks in the Barsky \& Sims sense: reallocative shocks, shocks to inventories, etc...  
\end{itemize}

@@

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{The current TFP slowdown}

\vspace{-1cm}
\noindent
\begin{figure}
\centering
\includegraphics[scale=0.3]{\ourTablePath fig_BS_fevtable}
\end{figure}

%%%%%%% Slide %%%%%%
\begin{frame}
	\frametitle{Production of IT goods}
	\label{it_sector}

Production function for IT goods:

\begin{equation}
V_{i,t} = \lambda_t \Psi_t f(S_{i,t}) 
\end{equation}
where $S_{i,t}$ is investment by producer $i$ in IT goods and $\lambda$ is the productivity of the IT sector.

@@

The IT producer's problem is to max discounted profits $J_{i,t+1}$

\begin{equation}
\max_{S_{i,t}} \; \mathbb{E}(\Lambda_{t,t+1},J_{i,t+1})\lambda_t \Psi_t f(S_{i,t}) - P^{IT}_t S_{i,t}
\end{equation}

\
FOC:

\begin{equation}
P^{IT}_t = \mathbb{E}(\Lambda_{t,t+1},J_{i,t+1})\lambda \Psi_t f_1
\end{equation}

\hyperlink{expanding_variety}{\beamerreturnbutton{Return}}	
\end{frame}
%%%%%%%%%%%%%%%%%
